---
title: "2010 and 2017 CBP Apprehensions Analysis"
author: "Janie Briglio, Ben Gaudiosi, Kayla Ippongi"
date: "2/14/2018"
output:
  pdf_document: default
  html_document: default
---
#Bar Graph Comparison of Border Partol Apprehensions 
entering the data for 2010 and making a column and row for the total for the locations 
and the months
```{r}
data1 <- read.csv("BP apprehensions 2010.csv", header = TRUE, stringsAsFactors = FALSE)
rownames(data1) <- data1[,1]
x <- subset(data1, select= -c(Sector))
x <- rbind(x, colSums(x))
rownames(x) <- c(rownames(x)[-length(rownames(x))], "Total")
x <- cbind(x,rowSums(x))
colnames(x) <- c(colnames(x)[-length(colnames(x))], "Total")
```
bar graph of the totals for for each location 
```{r}
barplot(x[1:9, 13], names.arg = rownames(x)[1:9], 
        las=2,
        axisnames=TRUE,
        main="2010 Border Patrol Apprehensions by Sector",
        border="black",
        col="blue")
```
entering the data for 2017 and making a column and row for the total for the locations 
and the months
```{r}
data2 <- read.csv("PB Apprehensions 2017.csv", header = TRUE, stringsAsFactors = FALSE)
data2 = head(data2[1:13],-1)
rownames(data2) <- data2[,1]
y <- subset(data2, select= -c(Sector))
y <- rbind(y, colSums(y))
rownames(y) <- c(rownames(y)[-length(rownames(y))], "Total")
y <- cbind(y, rowSums(y))
colnames(y) <- c(colnames(y)[-length(colnames(y))], "Total")
```
bar graph for the totals of each location 
```{r}
barplot(y[1:9, 13], names.arg = rownames(y)[1:9], 
        las=2,
        axisnames=TRUE,
        main="2017 Border Patrol Apprehensions by Sector",
        border="black",
        col="blue")
```

# Big Bend Comparison
\includegraphics{img/bigbend.png}

# Del Rio Comparison
\includegraphics{img/delrio.png}

# El Centrio Comparison
\includegraphics{img/elcentrio.png}

# El Paso Comparison
\includegraphics{img/elpaso.png}

# Laredo Comparison
\includegraphics{img/laredo.png}

# Rio Grande Valley Comparison
\includegraphics{img/riogrande.png}

# San Diego Comparison
\includegraphics{img/sandiego.png}

# Tuscon Comparison
\includegraphics{img/tuscon.png}

# Yuma Comparison
\includegraphics{img/yuma.png}

#Time Series
This time series graph represents the total yearly apprehensions from 2000-2017. The trend shows 
the total apprhensions slowly decreasing as the years move towards 2017.
```{r}
PBSum <- read.csv("PB monthly summaries.csv")
ts.plot(ts(as.vector(rev(t(rev(PBSum[-1])))), start = c(2000,10), frequency=12),
      ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[18,-1])),each=12)  )  )), start = c(2000,10), frequency=12),
      ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[17,-1])),each=12)  )  )), start = c(2001,10), frequency=12),
      ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[16,-1])),each=12)  )  )), start = c(2002,10), frequency=12),
      ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[15,-1])),each=12)  )  )), start = c(2003,10), frequency=12),
      ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[14,-1])),each=12)  )  )), start = c(2004,10), frequency=12),
      ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[13,-1])),each=12)  )  )), start = c(2005,10), frequency=12),
      ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[12,-1])),each=12)  )  )), start = c(2006,10), frequency=12),
      ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[11,-1])),each=12)  )  )), start = c(2007,10), frequency=12),
      ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[10,-1])),each=12)  )  )), start = c(2008,10), frequency=12),
      ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[9,-1])),each=12)  )  )), start = c(2009,10), frequency=12),
      ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[8,-1])),each=12)  )  )), start = c(2010,10), frequency=12),
      ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[7,-1])),each=12)  )  )), start = c(2011,10), frequency=12),
      ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[6,-1])),each=12)  )  )), start = c(2012,10), frequency=12),
      ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[5,-1])),each=12)  )  )), start = c(2013,10), frequency=12),
      ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[4,-1])),each=12)  )  )), start = c(2014,10), frequency=12),
      ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[3,-1])),each=12)  )  )), start = c(2015,10), frequency=12),
      ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[2,-1])),each=12)  )  )), start = c(2016,10), frequency=12),
      ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[1,-1])),each=12)  )  )), start = c(2017,10), frequency=12),
                                
      gpars=list(xlab="Year", ylab="Apprehensions", lty=1, col=c('blue',rep(c('red'),times=18), ltw=2))
                         );text(rev(PBSum[,1])+1,rev(c(rowMeans(PBSum[-1])))+4000,labels=paste(rev(PBSum[,1])),cex=0.6,col="red")

```

#TTests 
By looking at the resulting T-test values, March April and May have the highest T-test values which matches the 2010 CBP total apprehensions data - as March, April and May have the highest total apprensions in 2010 but the lowest in 2017. Comparatively, October November and December have the lowest T-test values, going into the negative range. This 3 month period has the highest total apprensions in 2017 but is not too different from it's 2010 data, which is why the t-test values are closer to zero.

entering data as a matrix and removing non numerical data 
```{r}
data1 <- read.csv("BP apprehensions 2010.csv", header = TRUE, stringsAsFactors = FALSE)
data2 <- read.csv("PB Apprehensions 2017.csv", header = TRUE, stringsAsFactors = FALSE)
data2 <- head(data2[1:13],-1)

x <- subset(data1, select= -c(Sector))
y <- subset(data2, select= -c(Sector))
```

running comparisons between the months for each location
##October Comparison:
```{r}
t.test(data1$October, data2$October, paired=TRUE)
```
##November Comparison:
```{r}
t.test(data1$November, data2$November, paired=TRUE)
```
##December Comparison:
```{r}
t.test(data1$December, data2$December, paired=TRUE)
```
##Janurary Comparison:
```{r}
t.test(data1$January, data2$January, paired=TRUE)
```
##Feburary Comparison: 
```{r}
t.test(data1$February, data2$February, paired=TRUE)
```
##March Comparison:
```{r}
t.test(data1$March, data2$March, paired=TRUE)
```
##April Comparison:
```{r}
t.test(data1$April, data2$April, paired=TRUE)
```
##May Comparison: 
```{r}
t.test(data1$May, data2$May, paired=TRUE)
```
##June Comparison:
```{r}
t.test(data1$June, data2$June, paired=TRUE)
```
##July Comparison:
```{r}
t.test(data1$July, data2$July, paired=TRUE)
```
##August Comparison:
```{r}
t.test(data1$August, data2$August, paired=TRUE)
```
##September Comparison: 
```{r}
t.test(data1$September, data2$September, paired=TRUE)
```

running a comparison between the totals of each location in both years 
```{r}
total2010<- rowSums(x[1:12])
total2017 <- rowSums(y[1:12])
t.test(total2010, total2017, paired=TRUE)
```
