---
title: "2010 and 2017 CBP Apprehensions Analysis"
author: "Janie Briglio, Ben Gaudiosi, Kayla Ippongi"
output: beamer_presentation
        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## 2010 Total Apprehensions  

```{r}
   BP2010 <- read.csv("BP apprehensions 2010.csv")
   rownames(BP2010) <- BP2010[,1]
   x <- subset(BP2010, select= -c(Sector))
   x <- rbind(x, colSums(x))
   rownames(x) <- c(rownames(x)[-length(rownames(x))], "Total")
   x <- cbind(x,rowSums(x))
   colnames(x) <- c(colnames(x)[-length(colnames(x))], "Total")
     barplot(x[1:9, 13], names.arg = rownames(x)[1:9], 
             las=2,
             axisnames=TRUE,
             main="2010 Total Apprehensions by Sector",
             border="black",
             col="red")
```

## 2010 Total Apprehensions: Analysis
- The data shows that Tuscon is sector with the highest total apphrensions in 2010 with 212,202 total
- On the other hand, Big Bend has the least apphrensions in 2010 with 5,288
                    

## 2017 Total Apprehensions 
```{r}
   PB2017 <- read.csv("PB Apprehensions 2017.csv")
   PB2017 = head(PB2017[1:13],-1)
   rownames(PB2017) <- PB2017[,1]
   y <- subset(PB2017, select= -c(Sector))
   y <- rbind(y, colSums(y))
   rownames(y) <- c(rownames(y)[-length(rownames(y))], "Total")
   y <- cbind(y, rowSums(y))
   colnames(y) <- c(colnames(y)[-length(colnames(y))], "Total")
     barplot(y[1:9, 13], names.arg = rownames(y)[1:9], 
             las=2,
             axisnames=TRUE,
             main="2017 Total Apprehensions by Sector",
             border="black",
             col="blue")  
```

## 2017 Total Apprehensions: Analysis
- This barplot identifies the summation of apprehensions in 2017 by each sector. There is a change in the maximum compared to 2010.
- This time Rio Grande Valley has the greatest total apprhensions, with 137,562 total (230% more than in 2010).
- While Tuscon had the max apprehensions in 2010, it has 38,657 this year, 81% less than 2010.


## Time Series

```{r}
      PBSum <- read.csv("PB monthly summaries.csv")
      ts.plot(ts(as.vector(rev(t(rev(PBSum[-1])))), start = c(2000,10), frequency=12),
                                ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[18,-1])),each=12)  )  )), start = c(2000,10), frequency=12),
                                ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[17,-1])),each=12)  )  )), start = c(2001,10), frequency=12),
                                ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[16,-1])),each=12)  )  )), start = c(2002,10), frequency=12),
                                ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[15,-1])),each=12)  )  )), start = c(2003,10), frequency=12),
                                ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[14,-1])),each=12)  )  )), start = c(2004,10), frequency=12),
                                ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[13,-1])),each=12)  )  )), start = c(2005,10), frequency=12),
                                ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[12,-1])),each=12)  )  )), start = c(2006,10), frequency=12),
                                ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[11,-1])),each=12)  )  )), start = c(2007,10), frequency=12),
                                ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[10,-1])),each=12)  )  )), start = c(2008,10), frequency=12),
                                ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[9,-1])),each=12)  )  )), start = c(2009,10), frequency=12),
                                ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[8,-1])),each=12)  )  )), start = c(2010,10), frequency=12),
                                ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[7,-1])),each=12)  )  )), start = c(2011,10), frequency=12),
                                ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[6,-1])),each=12)  )  )), start = c(2012,10), frequency=12),
                                ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[5,-1])),each=12)  )  )), start = c(2013,10), frequency=12),
                                ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[4,-1])),each=12)  )  )), start = c(2014,10), frequency=12),
                                ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[3,-1])),each=12)  )  )), start = c(2015,10), frequency=12),
                                ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[2,-1])),each=12)  )  )), start = c(2016,10), frequency=12),
                                ts(as.vector(  t(  rev(rep(c(rowMeans(PBSum[1,-1])),each=12)  )  )), start = c(2017,10), frequency=12),
                                
                                  gpars=list(xlab="Year", ylab="Apprehensions", 
                                                                      lty=1, col=c('blue',rep(c('red'),times=18), ltw=2))
                         );text(rev(PBSum[,1])+1,rev(c(rowMeans(PBSum[-1])))+4000,labels=paste(rev(PBSum[,1])),cex=0.6,col="red")
```




## Time Series: Analysis
- Looking at the data, there is a noticible but gradual trend of the total yearly apprehensions diminishing. 
- 2000 started out with the most apprehenisons while 2017 exhibited the lowest apprehensions

## 2010 and 2017 Sector Comparisons 
- the following 9 graphs represent a side by side comparison of each sector's apphrensions in the year 2010 and 2017
- El Centrio, San Diego and Tuscon exhibit a decrease in overall apprehensions from 2010
- El Paso, Big Bend and Yuma exhibit a slight increase in overall apprehnsions from 2010
- Del Rio and Laredo exhibit mixed results, with Del Rio spiking in the winter and falling in the summer and 
Laredo falling in the spring/summer and increasing during the winter

## Big Bend Comparison
\includegraphics{img/bigbend.png}

## Del Rio Comparison
\includegraphics{img/delrio.png}

## El Centrio Comparison
\includegraphics{img/elcentrio.png}

## El Paso Comparison
\includegraphics{img/elpaso.png}

## Laredo Comparison
\includegraphics{img/laredo.png}

## Rio Grande Valley Comparison
\includegraphics{img/riogrande.png}

## San Diego Comparison
\includegraphics{img/sandiego.png}

## Tuscon Comparison
\includegraphics{img/tuscon.png}

## Yuma Comparison
\includegraphics{img/yuma.png}

## T Test
- The following slides are the outputs of each comparison between the 2010 and 2017 respective months
- The data we're interested in is the t statistic outputed in the second row 
- By comparing all the t statitics from the following tests we can observe that March April and May have the highest T-test values
  which matches the 2010 CBP total apprehensions data. This makes sense because they have the highest T statistics. Meaning there is a 
  large gap between the apprhensions in 2010 and 2017 with those months 
- Comparatively, October November and December have the lowest T-test values, going into the negative range. This 3 month period has the highest total
  apprensions in 2017 but is not too different from it's 2010 data, which is why the t-test values are closer to zero.

## T Test: October Comparison
```{r}
     t.test(BP2010$October, PB2017$October, paired=TRUE)
```

## T Test: November Comparison
```{r}
    t.test(BP2010$November, PB2017$November, paired=TRUE)
```

## T Test: December Comparison
```{r}
    t.test(BP2010$December, PB2017$December, paired=TRUE)
```
## T Test: January Comparison
```{r}
    t.test(BP2010$January, PB2017$January, paired=TRUE)
```
## T Test: February Comparison
```{r}
    t.test(BP2010$February, PB2017$February, paired=TRUE)
```
## T Test: March Comparison
```{r}
    t.test(BP2010$March, PB2017$March, paired=TRUE)
```
## T Test: April Comparison
```{r}
    t.test(BP2010$April, PB2017$April, paired=TRUE)
```
## T Test: May Comparison
```{r}
    t.test(BP2010$May, PB2017$May, paired=TRUE)
```
## T Test: June Comparison
```{r}
    t.test(BP2010$June, PB2017$June, paired=TRUE)
```
## T Test: July Comparison
```{r}
    t.test(BP2010$July, PB2017$July, paired=TRUE)
```
## T Test: August Comparison
```{r}
    t.test(BP2010$August, PB2017$August, paired=TRUE)
```
## T Test: September Comparison
```{r}
    t.test(BP2010$September, PB2017$September, paired=TRUE)
```

## T Test: Total Comparison 
```{r}
x <- subset(BP2010, select= -c(Sector))
y <- subset(PB2017, select= -c(Sector))
total2010<- rowSums(x[1:12])
total2017 <- rowSums(y[1:12])
t.test(total2010, total2017, paired=TRUE)
```