---
title: "Stat Test Comparisons"
author: "Janie Briglio"
date: "2/14/2018"
output: html_document
---
entering data as a matrix and removing non numerical data 
```{r}
data1 <- read.csv("2010 data.csv", header = TRUE, stringsAsFactors = FALSE)
as.matrix(data1)
data2 <- read.csv("2017 data.csv", header = TRUE, stringsAsFactors = FALSE)
as.matrix(data2)
x <- subset(data1, select= -c(Sector))
y <- subset(data2, select= -c(Sector))
```

###T-tests
####by month
october: no significant difference 
```{r}
t.test(data1$October, data2$October, paired=TRUE)
```
november: no significant difference  
```{r}
t.test(data1$November, data2$November, paired=TRUE)
```
december: no significant difference 
```{r}
t.test(data1$December, data2$December, paired=TRUE)
```
january: no significant difference
```{r}
t.test(data1$January, data2$January, paired=TRUE)
```
february: no significant difference 
```{r}
t.test(data1$February, data2$February, paired=TRUE)
```
march: no significant difference 
```{r}
t.test(data1$March, data2$March, paired=TRUE)
```
april: no significant difference 
```{r}
t.test(data1$April, data2$April, paired=TRUE)
```
may: no significant difference
```{r}
t.test(data1$May, data2$May, paired=TRUE)
```
june: no significant difference
```{r}
t.test(data1$June, data2$June, paired=TRUE)
```
july: no significant difference
```{r}
t.test(data1$July, data2$July, paired=TRUE)
```
august: no significant difference
```{r}
t.test(data1$August, data2$August, paired=TRUE)
```
september: no significant difference 
```{r}
t.test(data1$September, data2$September, paired=TRUE)
```

####by totals of each location in both years 
```{r}
total2010<- rowSums(x[1:12])
total2017 <- rowSums(y[1:12])
t.test(total2010, total2017, paired=TRUE)
```
no significant difference